Details matter: Similar host galaxies can harbor different merging compact binary populations

### 2015-06-02 paper/paper.tex
more edits

\newcommand\mc{{\cal M}_c{}}
\newcommand\editremark[1]{{\color{red}#1}}
\newcommand\msun{M_\odot}
\newcommand\unit[1]{\text{#1}}
\newcommand\abbrvPSgrb{PSG}
\newcommand\abbrvPSellipticals{PSE}
\newcommand\abbrvKBLowZa{BD2010}
\newcommand\ForInternalReference[1]{}
\newcommand\modelDefault{A}
\newcommand\modelNiino{A'}
\newcommand\modelKW{D}
\newcommand\modelMannucci{D'}
\newcommand\modelKB{B}
\newcommand\modelPanter{C}
\newcommand\cqg{CQG}
\newcommand\E[1]{\left<#1\right>}
%\newcommand\aap{A\&A}
%\newcommand\apss{APSS}
%\newcommand\aaps{AAPS}
%% \newcommand\apjs{ApJ S}
%% \newcommand\aj{AJ}
%% \newcommand\apjl{ApJL}
%% \newcommand\mnras{MNRAS}
%% \newcommand\pasp{PASP}
%% \newcommand\pasj{PASJ}
%% \newcommand\araa{ARAA}
%% \newcommand\physrep{Phys. Rep.}

\begin{document}
\title{Details matter: Similar-looking host galaxies can produce different merging compact binary populations} 
\author{R.\ O'Shaughnessy}
\affiliation{Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, NY 14623, USA}
\email{oshaughn@mail.rit.edu}
\author{ J. Bellovary}
\affiliation{Department of Physics and Astronomy, Vanderbilt University, Nashville TN 37235}
\email{jillian.bellovary@vanderbilt.edu}
\begin{abstract}
% POINT:
% - existence proof by example .. not an attempt to solve the general problem we raise
Galaxy formation
simulations can reproduce present-day galaxies with notably different assembly and star formation histories.  
Recent binary evolution simulations suggest the compact binary merger rate depends
sensitively on the progentior's metallicity, to the extent that rare low-metallicity star formation during galaxy
assembly can produce more detected compact binaries than typical star formation.   
Using detailed simulations of galaxy and chemical evolution, we demonstrate by example that galaxies with similar present-day appearance can host distinctly different
compact binary populations.
We discuss the implications for transient multimessenger astronomy with compact binary sources. 
%% Present and future observations of  merging compact binaries can provide valuable constraints on their birthrate
%%  and formation scenarios.  However, recent binary evolution simulations suggest the compact binary merger rate depends
%% sensitively on the progentior's metallicity, to the extent that rare low-metallicity star formation during galaxy
%% assembly can produce more detected compact binaries than typical star formation.    Additionally, galaxy formation
%% simulations can reproduce present-day galaxies with notably different assembly and star formation histories, with
%% corresponding different present-day merger rates.  
%% % POINT 0: Gasoline + popsyn: real work was done
%% Using existing simulations of the assembly,  chemical evolution, and compact-object formation rate of a single halo
%% Milky Way-like galaxy with Gasoline, we construct present-day compact binary merger rates using a simple semianalytic
%% compact binary formation model ($dP/dt \propto Z^\alpha/t $) motivated by detailed calculations in the literature. 
%% % POINT 1: Observations
%% %   - difficult to reconstruct f
%% %   - local particle far away
%% As expected, we show the present-day galactic state can be only weakly correlated with the dominant compact binary
%% fomation rate: the galaxy could form in several ways, including epochs of low-metallicity star formation through major
%% mergers; and present-day mergers rarely occur near gas with their progenitor metallicity, even when not kicked.  
%% %  - but the galaxy 
%% By contrast,  detailed studies which reconstruct the galaxy's star formation history and metallicity evolution could,
%% when stacked, pinpoint the dominant formation process \editremark{need more than handwaving}.
% POINT 2: Revised merger rates
%% Averaging over the cosmological star formation history and galaxy mass-metallicity relation, we demonstrate low
%% metallicity star formation can produce much of the detected population of short GRBs and nearly all gravitatational-wave
%% sources. 
%% % POINT 3: Challenge for observations
%% Therefore,  though a potentially large fraction of binaries formed from low-metallicity gas,  their birth metallicity
%% cannot be identified directly, even with spatially resolved observations. 
%% %
%% We discuss the extent to which observations can distinguish between different scenarios for  low-metallicity compact
%% binary formation.
%
%\textbf{say something using GRBs in paragraph; perhaps mention relative bias in $p(M_g|GRB)$ relative to prior?}
%% \textbf{say something about different SFR histories making present galaxies -- can do more if we can distinguish history
%% of the galaxy}

@@

% POINT: Huge detail possible
The host galaxies of distant short GRBs have already been extensively investigated, with the associations being used to
draw preliminary conclusions about their progentitors  \cite{2014ARAA..52...43B}.   
%Of necessity, these analyses  on bulk correlations, often stacked to increase significance across multiple sources
%\editremark{clean up/don't piss off   Edo}.  
Unlike GRBs, detected gravitational wave sources will be limited by the range of LIGO to the local universe; for
example,  binary neutron star sources should be closer than $400\unit{Mpc}$.  Due to their proximity, each host galaxy
can be explored at great depth and detail via  position-resolved spectroscopy, enabling detailed position-resolved star-formation
and chemical evolution histories \editremark{citation: need better ones} \citep[see,e.g.][]{2009MNRAS.396..462K,2014MNRAS.444..336C}.
% POINT: But is it useful?
In this work, we demonstrate by concrete example that detailed analysis of individual galaxies can be essential to
investigate key physical questions about the origin of compact binary mergers.

@@

\noindent
where the star formation efficiency parameter $c^*$ is set to 0.1 such
that our galaxies match the observed Kennicutt-Schmidt law
\citep{1989ApJ...344..685K};  X$_{\rm H_2}$ is the molecular hydrogen fraction of the gas particle; $m_{star}$ and $m_{gas}$ are the star and gas
particle masses;\footnote{In our high-resolution simulations \textbf{ROS-check!}\editremark{this number is correct, is that what you want to check? - JMB}, each gas particle starts with a mass
  $m_{gas,0}=26,676 M_\odot$, gaining mass from feedback and losing it to star formation.  Each star particle, when
  formed, has $1/3$ of the progenitor mass of the forming gas particle.  See \cite{2010ApJ...717..121C} for a discussion
of resolution issues in these SPH simulations.} $t_{form}$ is the dynamical time for the gas
particle; and $\Delta t$ is the time between star formation episodes,
which we set to 1 Myr.  We model supernova feedback using the
blastwave formalism described in \citet{1977ApJ...218..148M} and implemented in
our simulations as in \citet{2006MNRAS.373.1074S}.  Each supernova releases
$E_{SN} = 10^{51}$ erg into the surrounding gas with a radius
determined by the blastwave equations.  These particles are not
allowed to cool for the duration of the blastwave, mimicking the
snowplow phase of a supernova explosion.  Previous works have found
that this set of parameters results in realistic galaxies which obey a
number of observed relations such as the mass-metallicity relation
\citep{2007ApJ...655L..17B}, the Tully-Fisher relation \citep{2009MNRAS.398..312G}, and
the size-luminosity relation \citep{2011ApJ...728...51B}, as well as reproduce
the detailed characteristics of bulgeless dwarf galaxies
\citep{2010Natur.463..203G} and the Milky Way \citep{2011ApJ...742...76G}.

@@

Binary evolution calculations suggest the detection rate depends sensitively on the assumed metallicity, in conjuction
with other parameters \citep[see,\,e.g.][and references
  therein]{popsyn-LowMetallicityImpact-Chris2008,popsyn-LIGO-SFR-2008}.\editremark{New dominik et al}
% POINT: 
%   - final BH spin, mass can change number favored
%   - of course
The detected population of compact binary mergers can further favor low-metallicity environments, e.g. through the total
mass of merging binaries.
% POINT: Simple model to encompass all
To explore plausible binary detection scenarios, we adopt a parameterized formalism for binary evolution and event detection, motivated by
the detailed studies of  \abbrvPSgrb{} and \abbrvPSellipticals.  Specifically, we encapsulate all information about
 binary evolution into two  parameterized response functions.  The funciton ${\cal G}(t| Z) =
dN/dt dM_*$ describes how many mergers per unit initial mass occur after a time $t$ since a starburst,
given ambient metallicity $Z$.  The function $p(m_1,m_2|Z)$ is a metallicity-dependent mass distribution for binaries
formed in a metallicity $Z$.
%
Finally, we account for the substantially increased sensitivity of gravitational wave detectors to binaries of larger
chirp mass $\mc=(m_1 m_2)^{3/5}/(m_1+m_2)^{1/5}$; see \abbrvPSellipticals.
%
In terms of this response function, we can therefore calculate the relative contribution to the LIGO  event rate $r_D$
due to any individual galaxy via the following formula 
\begin{widetext}
\begin{subequations}
\label{eq:RatePerGalaxy}
\begin{eqnarray}
r_D  &\equiv& \int^{0}_{-T_{univ}} dt' d\log Z  \, [\dot{M}_*(t') p(\log Z|t) ] \times \; {\cal G}(0-t' | Z)  
 [\int d\mc p(\mc|Z) (\mc/1.2 M_\odot)^{15/6}]
%\\
%% {\cal R}(z)&\equiv &  \int dx {\cal R} p(z)
\end{eqnarray}
\end{subequations}
\end{widetext}
In this expression, the galactic star formation rate
$\dot{M}_*$ and star formation metallicity distribution $p(\log Z|t)$ are galaxy-dependent and drawn from our
simulations.   \editremark{rewrite this to be average chirp mass at a given metallicity}

@@

\subsection{Tidal disruption and black hole mass versus metallicity}
At merger, black holes may promptly engulf a companion neutron star \editremark{same papers as for Brandon:Foucart,
  etc}, unless they have sufficiently small mass and sufficiently large spin.  Without  a tidal disruption event prior to merger, the prospects for associated electromagnetic counterparts
are severely diminished.  
%
A neutron star  (NS) coalescing with a black hole (BH) will be tidally disrupted prior to merger if and only if the component
masses, spins, and equation of state permit it.  As described below, only a small fraction of all  possible coalescing
BH-NS binaries have parameters consistent with tidal disruption.  
%
Stellar evolution and hence supernovae remnants are expected to evolve
substantially with birth metallicity, the black hole mass and possibly spin in black hole-neutron star binaries can depend sensitively on binary formation
channels and the fraction of coalescing NS-BH binaries which are tidally disrupted will vary
over cosmic time. 
%
To quantify this effect, we adopt a simple formula for tidal disruption of neutron stars in BH-NS binaries and a simple
phenomenological model for the metallicity-dependent mass distribution of BHs in BH-NS binaries

@@

% POINT: Trend versus mass and opening angle.  [Be careful: disruption time will depend on opening angle]
To estimate whether a specific compact binary will be tidally disrupted prior to merger, we employ the
following procecure to identify posterior samples consistent with tidal disruption of a neutron star.  First, the
smaller mass must lie within the range of masses allowed by our fiducial NS equation of state (\textbf{MPA1 from Ben}), here
between $0.5 M_\odot$ and $2.5 M_\odot$.  Second, the disruption process must leave behind a remnant disk with nonzero
mass, as estimated using the expression in \cite{2012PhRvD..86l4007F} (his Eqs. (6,12-13)):
\begin{align}
\frac{M_{disk}}{M_{NS}} &=
 0.288 (3 m_{bh}/m_{ns})^{1/3}[1-2 {\cal C}] 
\nonumber \\
&
 -0.148 (m_{bh}/m_{ns})  {\cal C} R_{isco}(a)/m_{bh} 
\label{eq:Mdisk}
\end{align}
where $q=m_{bh}/m_{NS}$ is the binary mass ratio; $a=S_{bh}/m_{bh}^2$ is a dimensionless measure of the black hole spin;
and  where ${\cal C}(m_{ns})=m_{ns}/R_{ns}$ is a mass- and equation-of-state dependent  measure of the neutron star
compactness.  In Foucart's expression, shown above, $R_{isco}(a)$ is the radius of the  innermost stable equitorial circular orbit  of a test particle
about a Kerr black hole  \citep{1972ApJ...178..347B}. % Brandon:  [their Eq. 2.21].   
When the black hole spin is not aligned with the orbit, we use the aligned component of the black hole spin magnitude $a\rightarrow a \cos \theta_{LS}$ in this
expression.   Fixing the neutron star mass $m_{ns}=1.4 M_\odot$ and hence compactness to ${\cal C}=1.66$, this expression implies a two-dimensional surface in
the three-dimensional parameter space $m_{bh},a,\theta_{LS}$ seperates parameters corresponding to tidal disruption to
parameters corresponding to prompt coalescence.
%
After coalescence, the remnant disk generates a tightly beamed electromagnetic counterpart.  For simplicity, we assume
beaming angle is independent of the disruption process or progenitor parameters, so a constant beaming correction factor $f_b$ enters into all
subsequent calculations and can hence be omitted.  
%
Figure \ref{fig:TidalDisrupt} shows contours of this expression for different values of $\theta_{LS}$.

>>>>

\newcommand\mc{{\cal M}_c{}}
\newcommand\editremark[1]{{\color{red}#1}}
\newcommand\jillianremark[1]{{\color{blue}#1}}
\newcommand\msun{M_\odot}
\newcommand\unit[1]{\text{#1}}
\newcommand\abbrvPSgrb{PSG}
\newcommand\abbrvPSellipticals{PSE}
\newcommand\abbrvKBLowZa{BD2010}
\newcommand\ForInternalReference[1]{}
\newcommand\modelDefault{A}
\newcommand\modelNiino{A'}
\newcommand\modelKW{D}
\newcommand\modelMannucci{D'}
\newcommand\modelKB{B}
\newcommand\modelPanter{C}
\newcommand\cqg{CQG}
\newcommand\E[1]{\left<#1\right>}
%\newcommand\aap{A\&A}
%\newcommand\apss{APSS}
%\newcommand\aaps{AAPS}
%% \newcommand\apjs{ApJ S}
%% \newcommand\aj{AJ}
%% \newcommand\apjl{ApJL}
%% \newcommand\mnras{MNRAS}
%% \newcommand\pasp{PASP}
%% \newcommand\pasj{PASJ}
%% \newcommand\araa{ARAA}
%% \newcommand\physrep{Phys. Rep.}

\begin{document}
\title{Details matter: Similar host galaxies can harbor different merging compact binary populations} 
\author{R.\ O'Shaughnessy}
\affiliation{Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, NY 14623, USA}
\email{oshaughn@mail.rit.edu}
\author{ J. Bellovary}
\affiliation{Department of Physics and Astronomy, Vanderbilt University, Nashville TN 37235}
\email{jillian.bellovary@vanderbilt.edu}
\begin{abstract}
% POINT:
% - existence proof by example .. not an attempt to solve the general problem we raise
Galaxy formation
simulations can reproduce present-day galaxies with notably different assembly and star formation histories.  
Recent simulations of compact binary evolution suggest the compact object merger rate depends
sensitively on the progentior's metallicity, to the extent that rare low-metallicity star formation during galaxy
assembly can produce more detected compact binaries than typical star formation.   
Using detailed simulations of galaxy and chemical evolution, we demonstrate by example that galaxies with similar present-day appearance can host distinctly different
compact binary populations.
We discuss the implications for transient multimessenger astronomy with compact binary sources. \jillianremark{Add quantifiable stuff when we have it.}
%% Present and future observations of  merging compact binaries can provide valuable constraints on their birthrate
%%  and formation scenarios.  However, recent binary evolution simulations suggest the compact binary merger rate depends
%% sensitively on the progentior's metallicity, to the extent that rare low-metallicity star formation during galaxy
%% assembly can produce more detected compact binaries than typical star formation.    Additionally, galaxy formation
%% simulations can reproduce present-day galaxies with notably different assembly and star formation histories, with
%% corresponding different present-day merger rates.  
%% % POINT 0: Gasoline + popsyn: real work was done
%% Using existing simulations of the assembly,  chemical evolution, and compact-object formation rate of a single halo
%% Milky Way-like galaxy with Gasoline, we construct present-day compact binary merger rates using a simple semianalytic
%% compact binary formation model ($dP/dt \propto Z^\alpha/t $) motivated by detailed calculations in the literature. 
%% % POINT 1: Observations
%% %   - difficult to reconstruct f
%% %   - local particle far away
%% As expected, we show the present-day galactic state can be only weakly correlated with the dominant compact binary
%% fomation rate: the galaxy could form in several ways, including epochs of low-metallicity star formation through major
%% mergers; and present-day mergers rarely occur near gas with their progenitor metallicity, even when not kicked.  
%% %  - but the galaxy 
%% By contrast,  detailed studies which reconstruct the galaxy's star formation history and metallicity evolution could,
%% when stacked, pinpoint the dominant formation process \editremark{need more than handwaving}.
% POINT 2: Revised merger rates
%% Averaging over the cosmological star formation history and galaxy mass-metallicity relation, we demonstrate low
%% metallicity star formation can produce much of the detected population of short GRBs and nearly all gravitatational-wave
%% sources. 
%% % POINT 3: Challenge for observations
%% Therefore,  though a potentially large fraction of binaries formed from low-metallicity gas,  their birth metallicity
%% cannot be identified directly, even with spatially resolved observations. 
%% %
%% We discuss the extent to which observations can distinguish between different scenarios for  low-metallicity compact
%% binary formation.
%
%\textbf{say something using GRBs in paragraph; perhaps mention relative bias in $p(M_g|GRB)$ relative to prior?}
%% \textbf{say something about different SFR histories making present galaxies -- can do more if we can distinguish history
%% of the galaxy}

@@

% POINT: Huge detail possible
The host galaxies of distant short GRBs have already been extensively investigated, with the associations being used to
draw preliminary conclusions about their progentitors  \cite{2014ARAA..52...43B}.   
%Of necessity, these analyses  on bulk correlations, often stacked to increase significance across multiple sources
%\editremark{clean up/don't piss off   Edo}.  
Unlike GRBs, detected gravitational wave sources will be limited by the range of LIGO to the local universe; for
example,  binary neutron star sources should be closer than $400\unit{Mpc}$.  Due to their proximity, each host galaxy
can be explored at great depth and detail via  position-resolved spectroscopy, enabling detailed position-resolved star-formation
and chemical evolution histories \editremark{citation: need better ones} \citep[see,e.g.][]{2009MNRAS.396..462K,2014MNRAS.444..336C}. \jillianremark{Do you want citations of GRB host observations or just general galaxy stuff?}
% POINT: But is it useful?
In this work, we demonstrate by concrete example that detailed analysis of individual galaxies can be essential to
investigate key physical questions about the origin of compact binary mergers.

@@

\noindent
where the star formation efficiency parameter $c^*$ is set to 0.1 such
that our galaxies match the observed Kennicutt-Schmidt law
\citep{1989ApJ...344..685K};  X$_{\rm H_2}$ is the molecular hydrogen fraction of the gas particle; $m_{star}$ and $m_{gas}$ are the star and gas
particle masses;\footnote{In our high-resolution simulations \textbf{ROS-check!}\jillianremark{this number is correct, is that what you want to check? - JMB}, each gas particle starts with a mass
  $m_{gas,0}=26,676 M_\odot$, gaining mass from feedback and losing it to star formation.  Each star particle, when
  formed, has $1/3$ of the progenitor mass of the forming gas particle.  See \cite{2010ApJ...717..121C} for a discussion
of resolution issues in these SPH simulations.} $t_{form}$ is the dynamical time for the gas
particle; and $\Delta t$ is the time between star formation episodes,
which we set to 1 Myr.  We model supernova feedback using the
blastwave formalism described in \citet{1977ApJ...218..148M} and implemented in
our simulations as in \citet{2006MNRAS.373.1074S}.  Each supernova releases
$E_{SN} = 10^{51}$ erg into the surrounding gas with a radius
determined by the blastwave equations.  These particles are not
allowed to cool for the duration of the blastwave, mimicking the
snowplow phase of a supernova explosion.  Previous works have found
that this set of parameters results in realistic galaxies which obey a
number of observed relations such as the mass-metallicity relation
\citep{2007ApJ...655L..17B}, the Tully-Fisher relation \citep{2009MNRAS.398..312G}, and
the size-luminosity relation \citep{2011ApJ...728...51B}, as well as reproduce
the detailed characteristics of bulgeless dwarf galaxies
\citep{2010Natur.463..203G} and the Milky Way \citep{2011ApJ...742...76G}.

@@

Binary evolution calculations suggest the detection rate depends sensitively on the assumed metallicity, in conjuction
with other parameters \citep[see,\,e.g.][and references
  therein]{popsyn-LowMetallicityImpact-Chris2008,popsyn-LIGO-SFR-2008}.\editremark{New dominik et al}
% POINT: 
%   - final BH spin, mass can change number favored
%   - of course
The detected population of compact binary mergers can further favor low-metallicity environments, e.g. through the total
mass of merging binaries. \jillianremark{The previous sentence is confusing - rephrase?  I have trouble with ``the detected population can further favor low-Z environments'' because that makes no sense.}
% POINT: Simple model to encompass all
To explore plausible binary detection scenarios, we adopt a parameterized formalism for binary evolution and event detection, motivated by
the detailed studies of  \abbrvPSgrb{} and \abbrvPSellipticals.  Specifically, we encapsulate all information about
 binary evolution into two  parameterized response functions.  The funciton ${\cal G}(t| Z) =
dN/dt dM_*$ describes how many mergers per unit initial mass occur after a time $t$ since a starburst,
given ambient metallicity $Z$.  The function $p(m_1,m_2|Z)$ is a metallicity-dependent mass distribution for binaries
formed in a metallicity $Z$.
%
Finally, we account for the substantially increased sensitivity of gravitational wave detectors to binaries of larger
chirp mass $\mc=(m_1 m_2)^{3/5}/(m_1+m_2)^{1/5}$; see \abbrvPSellipticals.
%
In terms of this response function, we can therefore calculate the relative contribution to the LIGO  event rate $r_D$
due to any individual galaxy via the following formula 
\begin{widetext}
\begin{subequations}
\label{eq:RatePerGalaxy}
\begin{eqnarray}
r_D  &\equiv& \int^{0}_{-T_{univ}} dt' d\log Z  \, [\dot{M}_*(t') p(\log Z|t) ] \times \; {\cal G}(0-t' | Z)  
 [\int d\mc p(\mc|Z) (\mc/1.2 M_\odot)^{15/6}]
%\\
%% {\cal R}(z)&\equiv &  \int dx {\cal R} p(z)
\end{eqnarray}
\end{subequations}
\end{widetext}
In this expression, the galactic star formation rate
$\dot{M}_*$ and star formation metallicity distribution $p(\log Z|t)$ are galaxy-dependent and drawn from our
simulations.   \editremark{rewrite this to be average chirp mass at a given metallicity}

@@

\subsection{Tidal disruption and black hole mass versus metallicity}
At merger, black holes may promptly engulf a companion neutron star \editremark{same papers as for Brandon:Foucart,
  etc}, unless they have sufficiently small mass and sufficiently large spin.  Without  a tidal disruption event prior to merger, the prospects for associated electromagnetic counterparts
are severely diminished.  
%
A neutron star  (NS) coalescing with a black hole (BH) will be tidally disrupted prior to merger if and only if the component
masses, spins, and equation of state permit it.  As described below, only a small fraction of all  possible coalescing
BH-NS binaries have parameters consistent with tidal disruption.  
%
\jillianremark{Grammatical nightmare ahead: }Stellar evolution and hence supernovae remnants are expected to evolve
substantially with birth metallicity, the black hole mass and possibly spin in black hole-neutron star binaries can depend sensitively on binary formation
channels and the fraction of coalescing NS-BH binaries which are tidally disrupted will vary
over cosmic time. 
%
To quantify this effect, we adopt a simple formula for tidal disruption of neutron stars in BH-NS binaries and a simple
phenomenological model for the metallicity-dependent mass distribution of BHs in BH-NS binaries

@@

% POINT: Trend versus mass and opening angle.  [Be careful: disruption time will depend on opening angle]
To estimate whether a specific compact binary will be tidally disrupted prior to merger, we employ the
following procecure to identify posterior samples consistent with tidal disruption of a neutron star.  First, the
smaller mass must lie within the range of masses allowed by our fiducial NS equation of state (\textbf{MPA1 from Ben}), here
between $0.5 M_\odot$ and $2.5 M_\odot$.  Second, the disruption process must leave behind a remnant disk with nonzero
mass, as estimated using the expression in \cite{2012PhRvD..86l4007F} (his Eqs. (6,12-13)):
\begin{align}
\frac{M_{disk}}{M_{NS}} &=
 0.288 (3 m_{bh}/m_{ns})^{1/3}[1-2 {\cal C}] 
\nonumber \\
&
 -0.148 (m_{bh}/m_{ns})  {\cal C} R_{isco}(a)/m_{bh} 
\label{eq:Mdisk}
\end{align}
where $q=m_{bh}/m_{NS}$ is the binary mass ratio; $a=S_{bh}/m_{bh}^2$ is a dimensionless measure of the black hole spin;
and  where ${\cal C}(m_{ns})=m_{ns}/R_{ns}$ is a mass- and equation-of-state dependent  measure of the neutron star
compactness.  In Foucart's expression, shown above, $R_{isco}(a)$ is the radius of the  innermost stable equitorial circular orbit  of a test particle
about a Kerr black hole  \citep{1972ApJ...178..347B}. % Brandon:  [their Eq. 2.21].   
When the black hole spin is not aligned with the orbit, we use the aligned component of the black hole spin magnitude $a\rightarrow a \cos \theta_{LS}$ in this
expression.   Fixing the neutron star mass $m_{ns}=1.4 M_\odot$ and hence compactness to ${\cal C}=1.66$, this expression implies a two-dimensional surface in
the three-dimensional parameter space $m_{bh},a,\theta_{LS}$ seperates parameters corresponding to tidal disruption to
parameters corresponding to prompt coalescence.
%
After coalescence, the remnant disk generates a tightly beamed electromagnetic counterpart.  For simplicity, we assume
beaming angle is independent of the disruption process or progenitor parameters, so a constant beaming correction factor $f_b$ enters into all
subsequent calculations and can hence be omitted.  
%
Figure \ref{fig:TidalDisrupt} shows contours of this expression for different values of $\theta_{LS}$.  \jillianremark{Add a sentence here describing the significance of the plot.  e.g. low-mass BHs can have a range of spins, regardless of angle, but high mass BHs always get high spins yay.}
